---
title: "Monte Carlo Confidence Intervals from Multiple Imputation Estimates"
author: Ivan Jacob Agaloos Pesigan
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Monte Carlo Confidence Intervals from Multiple Imputation Estimates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include = FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "fig-vignettes-",
  fig.cap = "",
  fig.width = 11,
  fig.height = 8,
  fig.retina = 2,
  dpi = 300,
  comment = "#>"
)
```

In this example, we use data from a study by Elliot and others (2007) on the effects of an intervention on healthy dietary behavior mediated by knowledge of healthy dietary behavior. This data was used as an empirical example in Yuan and MacKinnon (2009) and Wu and Jia (2013).

## Data

```{r}
#| message = FALSE
library(manMCMedMiss)
```

The data frame `elliot2007()`, which is included in the `manMCMedMiss` package, has 354 cases and 3 variables:

  - `x` - Intervention group membership.
  - `m` - Knowledge of healthy dietary behavior (knowledge post-intervention minus knowledge pre-intervention).
  - `y` - Healthy dietary behavior (behavior post-intervention minus behavior pre-intervention).

```{r}
knitr::kable(head(elliot2007))
```

## Amputation

Generate sample data with missing values. The missing data mechanism is missing at random (MAR). The proportion of missing cases is 0.30.

```{r}
set.seed(42)
data_missing <- AmputeData(
  elliot2007,
  mech = "MAR",
  prop = 0.10
)
knitr::kable(head(data_missing, n = 10))
```

## Multiple Imputation

Multiple data sets with complete data are generated from the original data set with missing values using multiple imputation.
There are several packages in `R` such as `Amelia`, `mice`, and `norm2` that can be used to generate completed data sets.
In the example below, we use the `mice` package to generate `m = 100` completed data sets using the joint specification approach
by specifying `method = "norm"`.
The data sets are saved in the list `mi` of length `m`.

```{r}
#| message = FALSE
library(mice)
mi <- complete(
  mice(
    data_missing,
    method = "norm",
    m = 100,
    print = FALSE,
    seed = 42
  ),
  action = "all"
)
```

## Model Fitting

We fit the model to each of the `m = 100` data sets using `lavaan`.
The output is saved in the list `fit`.

```{r}
#| message = FALSE
library(lavaan)
model <- "
  y ~ x + b * m
  m ~ a * x
"
fit <- lapply(
  X = mi,
  FUN = function(data,
                 model) {
    sem(
      model = model,
      data = data
    )
  },
  model = model
)
```

## Pooling Estimates

We pool the estimates to generate the estimated parameter vector (`param_vec`)
and the sampling variance-covariance matrix (`var_mat`).

We use the `lavaan::coef()` function to extract the parameter estimates
for each of the `m = 100` fitted models
and save the results in a list `param_vec`.

```{r}
param_vec <- lapply(
  X = fit,
  FUN = function(fit) {
    coef(fit)
  }
)
```

We use the `lavaan::vcov()` function to extract the sampling variance-covariance matrix
for each of the `m = 100` fitted models
and save the results in a list `var_mat`.

```{r}
var_mat <- lapply(
  X = fit,
  FUN = function(fit) {
    vcov(fit)
  }
)
```

The estimates are combined using the `manMCMedMiss::MICombine()` function.
The `manMCMedMiss::MICombine()` function uses a multivariate approach to combining
the entire sampling variance-covariance matrix and not just the diagonal elements.

```{r}
mi <- MICombine(
  param_vec = param_vec,
  var_mat = var_mat
)
```

A distribution of parameter estimates
are generated from the multivariate normal distribution
using the pooled parameter estimates vector
and the total sampling variance-covariance matrix.

```{r}
set.seed(42)
R <- 20000
mc <- MASS::mvrnorm(
  n = R,
  mu = mi$estimates,
  Sigma = mi$total
)
```

The corresponding quantiles for the 95% confidence intervals,
namely 2.5% and 97.5%,
are the lower and upper limits of the Monte Carlo confidence intervals.

```{r}
stats::quantile(
  mc[, "a"] * mc[, "b"],
  probs = c(0.025, 0.975)
)
```

