---
title: "Monte Carlo Method using Open-Source Software"
author: "Ivan Jacob Agaloos Pesigan"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Monte Carlo Method using Open-Source Software}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include = FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Stage 1: Estimation Stage

The estimation stage can be performed in your statistical package of choice. The parameter estimates and the sampling covariance matrix for the simple mediation model are given below.

### Parameter Estimates

```{r}
mu <- c(
  0.872,
  0.745
)
varnames <- c("alpha", "beta")
names(mu) <- varnames
mu
```

### Sampling Covariance Matrix

```{r}
Sigma <- matrix(
  data = c(
    1.253778e-02,
    5.244734e-06,
    5.244734e-06,
    1.429014e-02
  ),
  nrow = 2
)
rownames(Sigma) <- colnames(Sigma) <- varnames
Sigma
```

The parameter estimates of interest are $\hat{\alpha} = 0.872$ and $\hat{\beta} = 0.745$ and their corresponding sampling covariance matrix.

## Stage 2: Simulation Stage

In the simulation stage, $\hat{\alpha}$ and $\hat{\beta}$ are used as the mean vector $\mu$. The covariance matrix $\Sigma$ consists of the corresponding sampling covariance matrix.

### R

#### Simulation

The `MASS` package can be used in `R` to generate the sampling distribution of the indirect effect.

```{r}
library(MASS)
mustar <- mvrnorm(
  n = 20000,
  mu = mu,
  Sigma = Sigma
)
ab <- mustar[, 1] * mustar[, 2]
```

#### Plot

```{r}
hist(ab)
```

#### Confidence Intervals

The following corresponds to the $0.05$, $0.5$, $2.5$, $97.5$, $99.5$, $99.95$ percentiles.

```{r}
quantile(
  x = ab,
  probs = c(0.0005, 0.005, 0.025, 0.975, 0.995, 0.9995)
)
```

### Python

#### Simulation

The `Numpy` module can be used in `Python` to generate the sampling distribution of the indirect effect.

```python
import numpy as np
mu = [0.872, 0.745]
Sigma = [[1.253778e-02, 5.244734e-06], [5.244734e-06, 1.429014e-02]]
a, b = np.random.default_rng().multivariate_normal(mean = mu, cov = Sigma, size = 20000, method = 'eigh').T
ab = a * b
```

#### Confidence Intervals

The following corresponds to the $0.05$, $0.5$, $2.5$, $97.5$, $99.5$, $99.95$ percentiles.

```python
np.percentile(a = ab, q = [0.05, 0.5, 2.5, 97.5, 99.5, 99.95])
```

```python
array([0.27629, 0.34700399, 0.40509038, 0.92887276, 1.03758102, 1.15662989])
```

### Julia

#### Simulation

The `Distributions.jl` package can be used in `Julia` to generate the sampling distribution of the indirect effect.

```julia
using Random, Distributions
mustar = rand(MvNormal([0.872, 0.745], [[1.253778e-02, 5.244734e-06] [5.244734e-06, 1.429014e-02]]), 20000);
a = mustar[1, 1:20000];
b = mustar[2, 1:20000];
ab = a .* b;
```

#### Confidence Intervals

The following corresponds to the $0.05$, $0.5$, $2.5$, $97.5$, $99.5$, $99.95$ percentiles.

```julia
using StatsBase
percentile(ab, [0.05, 0.5, 2.5, 97.5, 99.5, 99.95])
```

```julia
6-element Vector{Float64}:
 0.2745307698592897
 0.34236235018811184
 0.4054190857337401
 0.9260797738723169
 1.0362695824493815
 1.16103686651973
```
