---
title: "Proportion of Missing Values"
author: Ivan Jacob Agaloos Pesigan
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Proportion of Missing Values}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include = FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "fig-vignettes-prop-",
  fig.width = 11,
  fig.height = 8,
  fig.retina = 2,
  dpi = 300,
  comment = "#>"
)
```

The following presents more details on the proportion of missing values used in the simulation study.

```{r}
#| message = FALSE
library(manMCMedMiss)
```

## Data Generation

```{r}
n <- 1000
tauprime <- 0
alpha <- 0.5
beta <- 0.5
```

```{r}
set.seed(42)
data_complete <- GenData(
  n = n,
  tauprime = tauprime,
  beta = beta,
  alpha = alpha
)
```

## Amputation

```{r}
prop <- 0.30
data_missing <- AmputeData(
  data_complete,
  mech = "MCAR",
  prop = prop
)
```

## Missing Data Patterns

The matrix below shows the missing data patterns in the amputed data.

```{r}
mice::md.pattern(data_missing)
```

### Proportion of missing values by row

This is the number of rows with missing values divided by the total number of rows. This is the value of the `prop` argument in `AmputeData()` which is a wrapper around the `mice::ampute()` function.

```{r}
sum_row <- sum(!complete.cases(data_missing))
prop_row <- sum_row / n
c(
  sum_row = sum_row,
  prop_row = prop_row
)
```

### Proportion of missing values by column

This is the number of rows per column with missing values divided by the total number of rows.

```{r}
sum_col <- colSums(is.na(data_missing))
names(sum_col) <- paste0(
  "sum_col_",
  names(sum_col)
)
prop_col <- sum_col / n
names(prop_col) <- gsub("sum", "prop", names(sum_col))
c(
  sum_col,
  prop_col
)
```

### Proportion of missing values by cell

This is the number of cells with missing values divided by the number of rows times the number of columns.

```{r}
sum_cell <- sum(is.na(data_missing))
prop_cell <- sum_cell / (n * 3)
prop_cell
c(
  sum_cell = sum_cell,
  prop_cell = prop_cell
)
```

## Empirical Proportion of Missing Values

Below is a simple simulation to show the empirical proportions for the combination of sample size `n` and `prop` used in the study for 5000 replications.

```{r}
#| echo = FALSE
bar <- function(mech) {
  foo <- function(x,
                  reps = 5000,
                  mech,
                  seed = 42) {
    set.seed(42)
    n <- x[1]
    prop <- x[2]
    data_complete <- GenData(n = n)
    colMeans(
      do.call(
        what = "rbind",
        args = lapply(
          X = 1:reps,
          FUN = function(i, mech) {
            data_missing <- AmputeData(
              data_complete,
              mech = mech,
              prop = prop
            )
            sum_row <- sum(!complete.cases(data_missing))
            prop_row <- sum_row / n
            row <- c(
              sum_row = sum_row,
              prop_row = prop_row
            )
            sum_col <- colSums(is.na(data_missing))
            names(sum_col) <- paste0(
              "sum_col_",
              names(sum_col)
            )
            prop_col <- sum_col / n
            names(prop_col) <- gsub("sum", "prop", names(sum_col))
            col <- c(
              sum_col,
              prop_col
            )
            sum_cell <- sum(is.na(data_missing))
            prop_cell <- sum_cell / (n * 3)
            prop_cell
            cell <- c(
              sum_cell = sum_cell,
              prop_cell = prop_cell
            )
            c(
              n = n,
              prop = prop,
              row,
              col,
              cell,
              reps = reps
            )
          },
          mech = mech
        )
      )
    )
  }
  do.call(
    what = "rbind",
    args = lapply(
      X = as.data.frame(
        t(
          expand.grid(
            n = unique(manMCMedMiss::params$n),
            prop = 1:3 * .10
          )
        )
      ),
      FUN = foo,
      mech = mech
    )
  )
}
root <- rprojroot::is_rstudio_project
prop_mar_file <- root$find_file(
  ".data-raw",
  "prop_mar.Rds"
)
prop_mcar_file <- root$find_file(
  ".data-raw",
  "prop_mcar.Rds"
)
if (!file.exists(prop_mar_file)) {
  prop_mar <- bar(mech = "MAR")
  saveRDS(
    object = prop_mar,
    file = prop_mar_file,
    compress = "xz"
  )
}
if (!file.exists(prop_mcar_file)) {
  prop_mcar <- bar(mech = "MCAR")
  saveRDS(
    object = prop_mcar,
    file = prop_mcar_file,
    compress = "xz"
  )
}
prop_mar <- readRDS(file = prop_mar_file)
prop_mcar <- readRDS(file = prop_mcar_file)
rownames(prop_mar) <- 1:(dim(prop_mar)[1])
rownames(prop_mcar) <- 1:(dim(prop_mcar)[1])
```

### MAR

```{r}
#| echo = FALSE
knitr::kable(prop_mar, digits = 2)
```

```{r}
#| echo = FALSE,
#| eval = FALSE
DT::formatRound(
  DT::datatable(
    prop_mar,
    options = list(
      scrollX = TRUE
    )
  ),
  columns = 3:12
)
```

### MCAR

```{r}
#| echo = FALSE
knitr::kable(prop_mcar, digits = 2)
```

```{r}
#| echo = FALSE,
#| eval = FALSE
DT::formatRound(
  DT::datatable(
    prop_mcar,
    options = list(
      scrollX = TRUE
    )
  ),
  columns = 3:12
)
```

<details>
<summary>
Code to perform the simulation.
</summary>
```r
bar <- function(mech) {
  foo <- function(x,
                  reps = 5000,
                  mech,
                  seed = 42) {
    set.seed(42)
    n <- x[1]
    prop <- x[2]
    data_complete <- GenData(n = n)
    colMeans(
      do.call(
        what = "rbind",
        args = lapply(
          X = 1:reps,
          FUN = function(i, mech) {
            data_missing <- AmputeData(
              data_complete,
              mech = mech,
              prop = prop
            )
            sum_row <- sum(!complete.cases(data_missing))
            prop_row <- sum_row / n
            row <- c(
              sum_row = sum_row,
              prop_row = prop_row
            )
            sum_col <- colSums(is.na(data_missing))
            names(sum_col) <- paste0(
              "sum_col_",
              names(sum_col)
            )
            prop_col <- sum_col / n
            names(prop_col) <- gsub("sum", "prop", names(sum_col))
            col <- c(
              sum_col,
              prop_col
            )
            sum_cell <- sum(is.na(data_missing))
            prop_cell <- sum_cell / (n * 3)
            prop_cell
            cell <- c(
              sum_cell = sum_cell,
              prop_cell = prop_cell
            )
            c(
              n = n,
              prop = prop,
              row,
              col,
              cell,
              reps = reps
            )
          },
          mech = mech
        )
      )
    )
  }
  do.call(
    what = "rbind",
    args = lapply(
      X = as.data.frame(
        t(
          expand.grid(
            n = unique(manMCMedMiss::params$n),
            prop = 1:3 * .10
          )
        )
      ),
      FUN = foo,
      mech = mech
    )
  )
}
prop_mar <- bar("MAR")
prop_mcar <- bar("MCAR")
```
</details>
