---
title: "semmcci: Monte Carlo Confidence Intervals from Multiple Imputation Estimates"
author: Ivan Jacob Agaloos Pesigan
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{semmcci: Monte Carlo Confidence Intervals from Multiple Imputation Estimates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include = FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "fig-vignettes-",
  fig.cap = "",
  fig.width = 11,
  fig.height = 8,
  fig.retina = 2,
  dpi = 300,
  comment = "#>"
)
```

```{r}
#| include = FALSE
if (!require("remotes")) install.packages("remotes")
remotes::install_version(
  package = "MASS",
  version = "7.3-58.1"
)
remotes::install_version(
  package = "mice",
  version = "3.14.0"
)
remotes::install_version(
  package = "party",
  version = "1.3-10"
)
remotes::install_version(
  package = "semmcci",
  version = "1.1.2"
)
```

In this example, we use the `semmcci` package to generate Monte Carlo confidence intervals from multiple imputation estimates as described in Pesigan & Cheung (2023). We use data from a study by Elliot and others (2007) on the effects of an intervention on healthy dietary behavior mediated by knowledge of healthy dietary behavior. This data was used as an empirical example in Yuan and MacKinnon (2009) and Wu and Jia (2013).

## Data

```{r}
#| message = FALSE
library(manMCMedMiss)
```

The data frame `elliot2007()`, which is included in the `manMCMedMiss` package, has 354 cases and 3 variables:

  - `x` - Intervention group membership.
  - `m` - Knowledge of healthy dietary behavior (knowledge post-intervention minus knowledge pre-intervention).
  - `y` - Healthy dietary behavior (behavior post-intervention minus behavior pre-intervention).

```{r}
knitr::kable(head(elliot2007))
```

## Amputation

Generate sample data with missing values. The missing data mechanism is missing at random (MAR). The proportion of missing cases is 0.30.

```{r}
set.seed(42)
data_missing <- AmputeData(
  elliot2007,
  mech = "MAR",
  prop = 0.10
)
knitr::kable(head(data_missing, n = 10))
```

## Multiple Imputation

Multiple data sets with complete data are generated from the original data set with missing values using multiple imputation.
There are several packages in `R` such as `Amelia`, and `mice` that can be used to generate completed data sets.
In the example below, we use the `mice` package to generate `m = 100` completed data sets using the joint specification approach
by specifying `method = "norm"`.

```{r}
#| message = FALSE
library(mice)
mi <- mice(
  data_missing,
  method = "norm",
  m = 100,
  print = FALSE,
  seed = 42
)
```

## Model Fitting

We fit the model using `lavaan`.
We do not need to deal with missing values in this stage.
The output is saved in the object `fit`.

```{r}
#| message = FALSE
library(lavaan)
model <- "
  y ~ x + b * m
  m ~ a * x
  x ~~ x
  indirect := a * b
"
fit <- sem(
  model = model,
  data = data_missing
)
```

## Monte Carlo Confidence Intervals

The `fit` `lavaan` object and `mi` `mids` object can then be passed to the `semmcci::MCMI()` function
to generate Monte Carlo confidence intervals
using multiple imputation.

```{r}
#| message = FALSE,
#| warning = FALSE
library(semmcci)
MCMI(fit, mi = mi, alpha = 0.05)
```

## References

Elliot, D. L., Goldberg, L., Kuehl, K. S., Moe, E. L., Breger, R. K., &
Pickering, M. A. (2007).
The PHLAME (Promoting Healthy Lifestyles: Alternative Models' Effects)
firefighter study: Outcomes of two models of behavior change.
*Journal of occupational and environmental medicine*, *49*(2), 204–213.
http://doi.org/10.1097/JOM.0b013e3180329a8d

Pesigan, I. J. A., & Cheung, S. F. (2023).
Monte Carlo confidence intervals for the indirect effect with missing data.
*Behavior Research Methods*.
https://doi.org/10.3758/s13428-023-02114-4

Wu, W., & Jia, F. (2013).
A new procedure to test mediation with missing data
through nonparametric bootstrapping and multiple imputation,
*Multivariate Behavioral Research*, *48*(5), 663-691.
http://doi.org/10.1080/00273171.2013.816235

Yuan, Y., & MacKinnon, D. P. (2009).
Bayesian mediation analysis.
*Psychological methods*, *14*(4), 301–322.
http://doi.org/10.1037/a0016972
